The \emph{FrobeniusThresholds} package for \emph{Macaulay2}

### 2019-06-08 Fsing/FThresholdsPaper/FThresholds.tex
minor changes to paper
Changes to paper: rerun examples

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : R = ZZ/11[x,y];

@@

o6 = 10
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i7 : nu(1, 0_R, J)

@@

o8 : InfiniteNumber
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i9 : R = ZZ/17[x,y,z];

@@

o12 = true
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : R = ZZ/17[x,y,z];

i2 : f = x^3 + y^4 + z^5; -- a diagonal polynomial

i3 : time nu(3, f)
     -- used 0.00552312 seconds

o3 = 3756

i4 : time nu(3, f, UseSpecialAlgorithms => false)
     -- used 0.155826 seconds

o4 = 3756
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : R = ZZ/5[x,y,z];

i2 : f = x^2*y^4 + y^2*z^7 + z^2*x^8;

i3 : time nu(5, f) -- uses binary search (default)
     -- used 0.313778 seconds

o3 = 1124

i4 : time nu(5, f, Search => Linear)
     -- used 0.441903 seconds

o4 = 1124

i5 : m = ideal(x, y, z);

o5 : Ideal of R

i6 : time nu(2, m, m^2) -- uses binary search (default)
     -- used 1.58258 seconds

o6 = 97

i7 : time nu(2, m, m^2, Search => Linear) -- but linear seach gets luckier
     -- used 0.531175 seconds

o7 = 97
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}
i8 : nu(5, f, ReturnList => true)

o8 = {0, 1, 8, 44, 224, 1124}

o8 : List

i9 : nu(5, f, Verbose => true)
nuInternal: using comparison test FrobeniusRoot
ν(1) = 0
ν(p^1) = 1
ν(p^2) = 8
ν(p^3) = 44
ν(p^4) = 224
ν(p^5) = 1124

o9 = 1124
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : R = ZZ/11[x,y,z];

i2 : f = x^3 + y^3 + z^3 + x*y*z;

i3 : time nu(3, f) -- ContainmentTest is set to FrobeniusRoot, by default
     -- used 0.116697 seconds

o3 = 1209

i4 : time nu(3, f, ContainmentTest => StandardPower)
     -- used 9.65013 seconds

o4 = 1209
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : R = ZZ/3[x,y];

i2 : m = ideal(x, y);

i3 : nu(3, m^5)

o3 = 10

i4 : nu(3, m^5, ContainmentTest => FrobeniusPower)

o4 = 8
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : R = ZZ/7[x,y];

i2 : f = (x-1)^3 - (y-2)^2;

i3 : nu(1, f)

o3 = infinity

o3 : InfiniteNumber

i4 : nu(1, f, IsLocal => false)

o4 = 5

o4 : QQ
\end{MyVerbatim}
}

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : R = ZZ/11[x, y, z]/(x^2 - y*(z - 1));

@@

o4 = false
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i49 : R = ZZ/13[x, y];

i50 : f = y*((y + 1) - (x - 1)^2)*(x - 2)*(x + y - 2);

i51 : isFJumpingExponent(3/4, f)

o51 = true
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : ZZ/5[x,y,z];

@@

      1  2
o3 = {-, -}
      3  5

o3 : List
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i4 : fpt(x^17 + y^20 + z^24) -- a diagonal polynomial

@@

i6 : ZZ/5[x,y];

@@

o7 : QQ
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i8 : fpt({x, y, x + y, x + 3*y}, {2, 6, 9, 10}) == oo

o8 = true
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i12 : f = x^2*(x + y)^3*(x + 3*y^2)^5;

i13 : fpt(f, Attempts => 0) -- a bad estimate

          1
o13 = {0, -}
          5

o13 : List

i14 : fpt(f, Attempts => 0, DepthOfSearch => 3) -- a better estimate

        21   22
o14 = {---, ---}
       124  125

@@

i15 : fpt(f, Attempts => 1, DepthOfSearch => 3) -- the right-hand

      endpoint (nu+1)/p^e is the F-pure threshold

       22
o15 = ---
      125

\end{MyVerbatim}
}

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i16 : f = x^6*y^4 + x^4*y^9 + (x^2 + y^3)^3;

i17 : fpt(f, Attempts => 1, DepthOfSearch => 3)

       17   7
o17 = {--, --}
       62  25

@@

i18 : fpt(f, Attempts => 2, DepthOfSearch => 3) -- the left-hand endpoint
      nu/(p^e-1) is the F-pure threshold

      17
o18 = --
      62

o18 : QQ

\end{MyVerbatim}
}

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i19 : f = x^3*y^11*(x + y)^8*(x^2 + y^3)^8;

i20 : fpt(f, DepthOfSearch => 3, Attempts => 2)

        3   7
o20 = {--, ---}
       62  125

o20 : List

i21 : fpt(f, DepthOfSearch => 3, Attempts => 4) -- more attempts *always*
      improve the estimate

        1   4
o21 = {--, --}
       20  75

@@

i22 : fpt(f, DepthOfSearch => 3, Attempts => 4, GuessStrategy => denominator)
      -- the exact answer, after a change of strategy

       1
o22 = --
      19

o22 : QQ
\end{MyVerbatim}
}
\medspace

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i23 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

i24 : fpt(f, DepthOfSearch => 3, Attempts => 5)

        3   1
o24 = {--, --}
       40  13

i25 : fpt(f, DepthOfSearch => 3, Attempts => 5, Bounds => oo)

        24   1
o25 = {---, --}
       313  13

o25 : List
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i26 : f = -2*x^10*y^5 - x^5*y^9 - 2*x^3*y^10 + 2*x^2*y^8 - 2*x*y^9;

i27 : numeric fpt(f, DepthOfSearch => 3)

o27 = {.153226, .158333}

o27 : List

i28 : numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)
      -- FinalAttempt improves the estimate slightly

o28 = {.153461, .158333}

o28 : List
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i29 : f = -2*x^10*y^5 - x^5*y^9 - 2*x^3*y^10 + 2*x^2*y^8 - 2*x*y^9;

i30 : timing numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)

o30 = {.153461, .158333}
      -- 1.29156 seconds

o30 : Time

i31 : timing numeric fpt(f, DepthOfSearch => 3, Attempts => 4) -- a better
      answer in less time

o31 = {.153333, .158333}
      -- .554321 seconds

o31 : Time

i32 : timing fpt(f, DepthOfSearch => 4) -- the exact answer in even less
      time

       98
o32 = ---
      625
      -- .274657 seconds

o32 : Time
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i33 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

i34 : fpt(f, DepthOfSearch => 3, FinalAttempt => true, Verbose => true)

@@

nu has been computed: nu = nu(3,f) = 9 ...

fpt lies in the interval [ nu/(p^e-1), (nu+1)/p^e ] = [ 9/124, 2/25 ] ...

@@

guessFPT narrowed the interval down to ( 3/40, 2/25 ) ...

@@

First F-signature computed: s(f,(nu-1)/p^e) = 456/15625 ...

Second F-signature computed: s(f,nu/p^e) = 64/15625 ...

@@

fpt lies in the interval ( 3/40, 2/25 ).

        3   2
o34 = {--, --}
       40  25

o34 : List
\end{MyVerbatim}
}
\medspace

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : R = ZZ/7[x,y];

i2 : f = (y-1)^2-(x-1)^3;

i3 : fpt(f, IsLocal => false)

     5
o3 = -
     6

o3 : QQ

i4 : fpt(f)

o4 = infinity

o4 : InfiniteNumber
\end{MyVerbatim}
}
\medspace
% R = ZZ/7[x,y];
% f = (y-1)^2-(x-1)^3;
% fpt(f, IsLocal => false)
% fpt(f)

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : R = ZZ/7[x, y, z];

i2 : isSimpleNormalCrossing(x^3*y^2)

o2 = true

i3 : isSimpleNormalCrossing(x^2-y^2)

i4 : isSimpleNormalCrossing(x*y*(x-y))

o4 = false

i5 : isSimpleNormalCrossing(x^2-y*z)

o5 = false
\end{MyVerbatim}
}
\medspace

%             R = ZZ/7[x,y,z];
%             isSimpleNormalCrossing(x^3*y^2)
%             isSimpleNormalCrossing(x^2-y^2)
%             isSimpleNormalCrossing(x*y*(x-y))
%             isSimpleNormalCrossing(x^2-y*z)

@@

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}

i1 : R = QQ[x, y, z];

i2 : f = (y - (x-1)^2)*y^2; --SNC at the origin, but not globally

i3 : isSimpleNormalCrossing(f)

i4 : isSimpleNormalCrossing(f, IsLocal=>false)

o4 = false

i5 : g = (y-1)^2+(x-7)^2 --doesn't even pass through the origin

       2    2
o5 = x  + y  - 14x - 2y + 50

o5 : R

i6 : isSimpleNormalCrossing(g)

o6 = true

i7 : isSimpleNormalCrossing(g, IsLocal=>false)

o7 = false

i8 : h = x*y^2*(x+1)^3*(y-1)^4; --SNC everywhere

i9 : isSimpleNormalCrossing(h)

o9 = true

i10 : isSimpleNormalCrossing(h, IsLocal=>false)

o10 = true
\end{MyVerbatim}
}
\medspace

@@

%             g = (y-1)^2+(x-7)^2 --doesn't even pass through the origin

@@

%             h = x*y^2*(x+1)^3*(y-1)^4; --SNC everywhere

%             isSimpleNormalCrossing(h)

%             isSimpleNormalCrossing(h, IsLocal=>false)

>>>>

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/11[x,y];

@@

o6 = 10
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i7 : nu(1, 0_R, J)

@@

o8 : InfiniteNumber
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i9 : R = ZZ/17[x,y,z];

@@

o12 = true
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i13 : R = ZZ/17[x,y,z];

i14 : f = x^3 + y^4 + z^5; -- a diagonal polynomial

i15 : time nu(3, f)
     -- used 0.0108065 seconds

o15 = 3756

i16 : time nu(3, f, UseSpecialAlgorithms => false)
     -- used 0.209325 seconds

o16 = 3756
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i17 : R = ZZ/5[x,y,z];

i18 : f = x^2*y^4 + y^2*z^7 + z^2*x^8;

i19 : time nu(5, f) -- uses binary search (default)
     -- used 0.350179 seconds

o19 = 1124

i20 : time nu(5, f, Search => Linear)
     -- used 0.485823 seconds

o20 = 1124

i21 : m = ideal(x, y, z);

o21 : Ideal of R

i22 : time nu(2, m, m^2) -- uses binary search (default)
     -- used 1.71206 seconds

o22 = 97

i23 : time nu(2, m, m^2, Search => Linear) -- but linear seach gets luckier
     -- used 0.536135 seconds

o23 = 97
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i24 : nu(5, f, ReturnList => true)

o24 = {0, 1, 8, 44, 224, 1124}

i25 : nu(5, f, Verbose => true)

nuInternal: using comparison test FrobeniusRoot

ν(1) = 0
ν(p^1) = 1
ν(p^2) = 8
ν(p^3) = 44
ν(p^4) = 224
ν(p^5) = 1124

o25 = 1124
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i26 : R = ZZ/11[x,y,z];

i27 : f = x^3 + y^3 + z^3 + x*y*z;

i28 : time nu(3, f) -- ContainmentTest is set to FrobeniusRoot, by default
     -- used 0.127375 seconds

o28 = 1209

i29 : time nu(3, f, ContainmentTest => StandardPower)
     -- used 9.57149 seconds

o29 = 1209
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i30 : R = ZZ/3[x,y];

i31 : m = ideal(x, y);

o31 : Ideal of R

i32 : nu(3, m^5)

o32 = 10

i33 : nu(3, m^5, ContainmentTest => FrobeniusPower)

o33 = 8
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i34 : R = ZZ/7[x,y];

i35 : f = (x - 1)^3 - (y - 2)^2;

i36 : nu(1, f)

o36 = infinity

o36 : InfiniteNumber

i37 : nu(1, f, IsLocal => false)

o37 = 5
\end{MyVerbatim}
}

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/11[x, y, z]/(x^2 - y*(z - 1));

@@

o4 = false
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/11[x, y, z]/(x^2 - y*(z - 1));

o4 = false
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/5[x,y,z];

@@

       7  2
o3 = {--, -}
      25  5

o3 : List
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i4 : fpt(x^17 + y^20 + z^24) -- a diagonal polynomial

@@

i6 : R = ZZ/5[x,y];

@@

o7 : QQ
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i8 : fpt({x, y, x + y, x + 3*y}, {2, 6, 9, 10}) == oo

o8 = true
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i9 : f = x^2*(x + y)^3*(x + 3*y^2)^5;

i10 : fpt(f, Attempts => 0) -- a bad estimate

          1
o10 = {0, -}
          5

o10 : List

i11 : fpt(f, Attempts => 0, DepthOfSearch => 3) -- a better estimate

        21   22
o11 = {---, ---}
       124  125

o11 : List

i12 : fpt(f, Attempts => 1, DepthOfSearch => 3) -- the right-hand
      endpoint (nu+1)/p^e is the F-pure threshold

       22
o12 = ---
      125

o12 : QQ
\end{MyVerbatim}
}

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i13 : f = x^6*y^4 + x^4*y^9 + (x^2 + y^3)^3;

i14 : fpt(f, Attempts => 1, DepthOfSearch => 3)

       17   7
o14 = {--, --}
       62  25

@@

i15 : fpt(f, Attempts => 2, DepthOfSearch => 3) -- the left-hand endpoint
      nu/(p^e-1) is the F-pure threshold

      17
o15 = --
      62

o15 : QQ
\end{MyVerbatim}
}

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i16 : f = x^3*y^11*(x + y)^8*(x^2 + y^3)^8;

i17 : fpt(f, DepthOfSearch => 3, Attempts => 2)

        3   7
o17 = {--, ---}
       62  125

@@

i18 : fpt(f, DepthOfSearch => 3, Attempts => 4) -- more attempts *always*
      improve the estimate

        1   4
o18 = {--, --}
       20  75

o18 : List

i19 : fpt(f, DepthOfSearch => 3, Attempts => 4, GuessStrategy => denominator)
      -- the exact answer, after a change of strategy

       1
o19 = --
      19

o19 : QQ
\end{MyVerbatim}
}
\bigskip

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i20 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

i21 : fpt(f, DepthOfSearch => 3, Attempts => 5)

        3   1
o21 = {--, --}
       40  13

@@

i22 : fpt(f, DepthOfSearch => 3, Attempts => 5, Bounds => oo)

        24   1
o22 = {---, --}
       313  13

o22 : List
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i23 : f = -2*x^10*y^5 - x^5*y^9 - 2*x^3*y^10 + 2*x^2*y^8 - 2*x*y^9;

i24 : numeric fpt(f, DepthOfSearch => 3)

o24 = {.153226, .158333}

i25 : numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)
      -- FinalAttempt improves the estimate slightly

o25 = {.153461, .158333}

o25 : List
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i26 : f = -2*x^10*y^5 - x^5*y^9 - 2*x^3*y^10 + 2*x^2*y^8 - 2*x*y^9;

i27 : timing numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)

o27 = {.153461, .158333}
      -- 1.37902 seconds

o27 : Time

i28 : timing numeric fpt(f, DepthOfSearch => 3, Attempts => 4) -- a better
      answer in less time

o28 = {.153333, .158333}
      -- .614248 seconds

o28 : Time

i29 : timing fpt(f, DepthOfSearch => 4) -- the exact answer in even less
      time

       98
o29 = ---
      625
      -- .297493 seconds

o29 : Time
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i30 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

i31 : fpt(f, DepthOfSearch => 3, FinalAttempt => true, Verbose => true)

@@

ν has been computed: ν = ν(3,f) = 9 ...

fpt lies in the interval [ν/(p^e-1),(ν+1)/p^e] = [9/124,2/25] ...

@@

guessFPT narrowed the interval down to (3/40,2/25) ...

@@

First F-signature computed: s(f,(ν-1)/p^e) = 456/15625 ...

Second F-signature computed: s(f,ν/p^e) = 64/15625 ...

@@

fpt lies in the interval (3/40,2/25).

        3   2
o31 = {--, --}
       40  25

o31 : List
\end{MyVerbatim}
}
\bigskip

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i32 : R = ZZ/7[x,y];

i33 : f = (y - 1)^2-(x - 1)^3;

i34 : fpt(f, IsLocal => false)

      5
o34 = -
      6

o34 : QQ

i35 : fpt(f)

o35 = infinity

o35 : InfiniteNumber
\end{MyVerbatim}
}
\bigskip
% R = ZZ/7[x,y];
% f = (y-1)^2-(x-1)^3;
% fpt(f, IsLocal => false)
% fpt(f)

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i36 : R = ZZ/7[x, y, z];

i37 : isSimpleNormalCrossing(x^2 - y^2)

o37 = true

i38 : isSimpleNormalCrossing(x^2 - y*z)

o38 = false
\end{MyVerbatim}
}
\bigskip

%             R = ZZ/7[x,y,z];
%             isSimpleNormalCrossing(x^2-y^2)
%             isSimpleNormalCrossing(x^2-y*z)

@@

\bigskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i39 : R = QQ[x, y, z];

i40 : f = (y - (x - 1)^2)*y^2; --SNC at the origin, but not globally

i41 : isSimpleNormalCrossing(f)

o41 = true

i42 : isSimpleNormalCrossing(f, IsLocal => false)

o42 = false

i43 : g = x*y^2*(x + 1)^3*(y - 1)^4; --SNC everywhere

i44 : isSimpleNormalCrossing(g)

o44 = true

i45 : isSimpleNormalCrossing(g, IsLocal => false)

o45 = true
\end{MyVerbatim}
}
\bigskip

@@

%             g = x*y^2*(x+1)^3*(y-1)^4; --SNC everywhere

@@
